---
title: "Summary of Tier 2 NH3 Emission Factor Method"
author: "Chih-Yu Hung"
format:
  html:
    toc: true
    toc-depth: 2
    code-fold: true

---

## Overview

This report summarizes the proposed Tier 2 ammonia (`NH3-N`) emission factor
outputs for animal housing and manure storage. 

Housing emission factors are reported on a Total N basis and storage emission
factors are reported on a TAN basis. Figures therefore compare patterns and
relative magnitudes, but the denominator differs between the two stages.



## Preparation

Read the results from housing and storage and MC results. 

Note that we only have 4 board climate for housing, while 9 climate zones for storage. 

There are 84 EFs for housing (21 animal/housing/manure combination x 4 climates) and 288 EFs for storage (8 animal/stroage/manure combination x 4 removal scenarios x 9 climates)

```{r}
#| label: preparation
#| message: false
#| warning: false

library(dplyr)
library(ggplot2)
library(knitr)
library(readr)
library(tidyr)

paths <- list(
  housing_csv = file.path("Outputs", "Proposed_Tier2_Housing.csv"),
  housing_mc_csv = file.path("Outputs", "Proposed_Tier2_Housing_MC_Summary.csv"),
  storage_csv = file.path("Outputs", "Proposed_Tier2_Storage.csv"),
  storage_mc_csv = file.path("Outputs", "Proposed_Tier2_Storage_MC_Summary.csv"),
  figure_dir = file.path("Outputs", "Summary_Tier2_Method_Figures")
)

dir.create(paths$figure_dir, showWarnings = FALSE, recursive = TRUE)

housing_climate_order <- c(
  "Tropical",
  "Warm_Temperate",
  "Cool_Temperate",
  "Boreal"
)

storage_climate_order <- c(
  "Tropical Wet",
  "Tropical Moist",
  "Tropical Dry",
  "Warm Temperate Moist",
  "Warm Temperate Dry",
  "Cool Temperate Moist",
  "Cool Temperate Dry",
  "Boreal Moist",
  "Boreal Dry"
)

removal_schedule_order <- c(
  "April",
  "October",
  "April_October",
  "April_August_November"
)

animal_order <- c(
  "Dairy Cattle",
  "Non-Dairy Cattle",
  "Layers",
  "Broilers",
  "Goat",
  "Sheep",
  "Swine",
  "Swine_Finising",
  "Swine_Sows and Piglets"
)

ef_fill_scale <- scale_fill_gradientn(
  colours = c("#d8ead2", "#f0df9d", "#d69a45", "#a94438"),
  limits = c(0, 1),
  name = "EF"
)

format_ef <- function(value) {
  sprintf("%.3f", value)
}

save_plot <- function(plot, file_name, width, height) {
  ggplot2::ggsave(
    filename = file.path(paths$figure_dir, file_name),
    plot = plot,
    width = width,
    height = height,
    dpi = 300,
    bg = "white"
  )
}

theme_tier2 <- function(base_size = 10) {
  theme_minimal(base_size = base_size) +
    theme(
      panel.grid = element_blank(),
      plot.title = element_text(face = "bold"),
      plot.subtitle = element_text(color = "#4f5963"),
      axis.title = element_text(face = "bold"),
      strip.text = element_text(face = "bold"),
      legend.position = "bottom"
    )
}
```

## Read Outputs

```{r}
#| label: read-outputs
#| message: false
#| warning: false

housing <- readr::read_csv(paths$housing_csv, show_col_types = FALSE)
housing_mc <- readr::read_csv(paths$housing_mc_csv, show_col_types = FALSE)
storage <- readr::read_csv(paths$storage_csv, show_col_types = FALSE)
storage_mc <- readr::read_csv(paths$storage_mc_csv, show_col_types = FALSE)

housing_long <- housing %>%
  mutate(Animal_Type = factor(Animal_Type, levels = animal_order)) %>%
  tidyr::pivot_longer(
    cols = all_of(housing_climate_order),
    names_to = "Climate_zone",
    values_to = "EF_Deterministic"
  ) %>%
  mutate(
    Stage = "Housing",
    Row_Label = paste(Animal_Type, Housing, Manure_type, sep = " | "),
    Climate_zone = factor(Climate_zone, levels = housing_climate_order)
  )

housing_mc_plot <- housing_mc %>%
  mutate(
    Animal_Type = factor(Animal_Type, levels = animal_order),
    Stage = "Housing",
    Row_Label = paste(Animal_Type, Housing, Manure_type, Climate_zone, sep = " | "),
    Climate_zone = factor(Climate_zone, levels = housing_climate_order),
    MC_Width = EF_MC_Upper95 - EF_MC_Lower95
  )

storage_plot <- storage %>%
  mutate(
    Animal_Type = factor(Animal_Type, levels = animal_order),
    Stage = "Storage",
    Row_Label = paste(Animal_Type, Storage, Manure_type, sep = " | "),
    Climate_zone = factor(Climate_zone, levels = storage_climate_order),
    Removal_Schedule = factor(Removal_Schedule, levels = removal_schedule_order)
  )

storage_mc_plot <- storage_mc %>%
  mutate(
    Animal_Type = factor(Animal_Type, levels = animal_order),
    Stage = "Storage",
    Row_Label = paste(
      Animal_Type,
      Storage,
      Manure_type,
      Removal_Schedule,
      Climate_zone,
      sep = " | "
    ),
    Climate_zone = factor(Climate_zone, levels = storage_climate_order),
    Removal_Schedule = factor(Removal_Schedule, levels = removal_schedule_order),
    MC_Width = EF_MC_Upper95 - EF_MC_Lower95
  )
```


## Figure 1. Housing EF Heatmap

The results used DATAMAN EFs as the reference EF and assumed the EFs were obtained at 15 C. CTs of Tied housing and deep litter were assumed same to Feedlot; CT of Layers in aviary house was assumed same to broiler; CTs of sows and pigelets were assumed same to finisher. 

The housing heatmap shows a strong climate-temperature gradient, with highest
values in the tropical zone and lowest values in the boreal zone.

```{r}
#| label: fig-housing-heatmap
#| fig-width: 9
#| fig-height: 7

housing_heatmap_data <- housing_long %>%
  arrange(Animal_Type, Housing, Manure_type) %>%
  mutate(Row_Label = factor(Row_Label, levels = rev(unique(Row_Label))))

fig_housing_heatmap <- ggplot(
  housing_heatmap_data,
  aes(x = Climate_zone, y = Row_Label, fill = EF_Deterministic)
) +
  geom_tile(color = "white", linewidth = 0.4) +
  geom_text(aes(label = format_ef(EF_Deterministic)), size = 2.6) +
  ef_fill_scale +
  labs(
    title = "Housing deterministic NH3 emission factors",
    subtitle = "Total-N basis; four broad climate zones",
    x = "Climate zone",
    y = NULL
  ) +
  theme_tier2(base_size = 9) +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))

fig_housing_heatmap
save_plot(fig_housing_heatmap, "fig_01_housing_heatmap.png", 9, 7)
```



## Figure 2. Storage EF Heatmaps By Removal Schedule

Storage EFs are highest in warm climates and under less frequent removal,
because TAN can accumulate and remain available during high-volatilization
months. There is no CT for swine lagoon.

I found NH3 EF is not sensitive to removal frequency, unlike MCF. For instance,
Dairy/manure heap/solid in tropical wet has 0.289 at single Oct removal, while
0.276 with three removals. I think it is because of high corrected EF that
volatilized abundance of TAN before removal.

The annual NH3 EFs from storage was calculated as followed:

EF = Annual NH3 emissions / Total avaiable TAN in storage.

```{r}
#| label: fig-storage-heatmaps
#| fig-width: 13
#| fig-height: 8

storage_heatmap_data <- storage_plot %>%
  arrange(Animal_Type, Storage, Manure_type) %>%
  mutate(Row_Label = factor(Row_Label, levels = rev(unique(Row_Label))))

fig_storage_heatmaps <- ggplot(
  storage_heatmap_data,
  aes(x = Climate_zone, y = Row_Label, fill = Year3_EF_TAN)
) +
  geom_tile(color = "white", linewidth = 0.35) +
  geom_text(aes(label = format_ef(Year3_EF_TAN)), size = 2.1) +
  facet_wrap(~ Removal_Schedule, ncol = 2) +
  ef_fill_scale +
  labs(
    title = "Storage deterministic year-3 NH3 emission factors",
    subtitle = "TAN-available-weighted year-3 monthly NH3-N / available TAN",
    x = "Climate zone",
    y = NULL
  ) +
  theme_tier2(base_size = 8) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.spacing = unit(1, "lines")
  )

fig_storage_heatmaps
save_plot(fig_storage_heatmaps, "fig_02_storage_heatmaps.png", 13, 8)
```



## Figure 3. Monte Carlo Uncertainty Intervals

MC results showed the difference between deterministic EF and the mean of MC < 0.001 for housing and <0.01 for storage. However, wide ranges were observed in warm temperate and tropical regions as the exponential increase in EF under warm temperature. 


```{r}
#| label: fig-mc-housing
#| fig-width: 10
#| fig-height: 7

top_housing_uncertainty <- housing_mc_plot %>%
  arrange(desc(EF_MC_Mean)) %>%
  slice_head(n = 20) %>%
  arrange(Animal_Type, Housing, Manure_type, Climate_zone) %>%
  mutate(Row_Label = factor(Row_Label, levels = rev(Row_Label)))

fig_mc_housing <- ggplot(
  top_housing_uncertainty,
  aes(x = EF_MC_Mean, y = Row_Label)
) +
  geom_errorbar(
    aes(xmin = EF_MC_Lower95, xmax = EF_MC_Upper95),
    orientation = "y",
    width = 0.22,
    color = "#69737d"
  ) +
  geom_point(aes(color = EF_Deterministic), size = 2.4) +
  scale_color_gradient(low = "#5f9f6e", high = "#a94438", name = "Deterministic EF") +
  labs(
    title = "Housing Monte Carlo uncertainty intervals",
    subtitle = "Top 20 combinations by 95% interval width, points represent the average of EF",
    x = "EF, kg NH3-N / kg Total N",
    y = NULL
  ) +
  theme_tier2(base_size = 9)

fig_mc_housing
save_plot(fig_mc_housing, "fig_03a_housing_mc_intervals.png", 10, 7)
```

```{r}
#| label: fig-mc-storage
#| fig-width: 10
#| fig-height: 7

top_storage_uncertainty <- storage_mc_plot %>%
  arrange(desc(EF_MC_Mean)) %>%
  slice_head(n = 20) %>%
  arrange(Animal_Type, Storage, Manure_type, Removal_Schedule, Climate_zone) %>%
  mutate(Row_Label = factor(Row_Label, levels = rev(Row_Label)))

fig_mc_storage <- ggplot(
  top_storage_uncertainty,
  aes(x = EF_MC_Mean, y = Row_Label)
) +
  geom_errorbar(
    aes(xmin = EF_MC_Lower95, xmax = EF_MC_Upper95),
    orientation = "y",
    width = 0.22,
    color = "#69737d"
  ) +
  geom_point(aes(color = EF_Deterministic), size = 2.4) +
  scale_color_gradient(low = "#5f9f6e", high = "#a94438", name = "Deterministic EF") +
  labs(
    title = "Storage Monte Carlo uncertainty intervals",
    subtitle = "Top 20 combinations by 95% interval width, displayed in source animal order",
    x = "EF, kg NH3-N / kg TAN",
    y = NULL
  ) +
  theme_tier2(base_size = 9)

fig_mc_storage
save_plot(fig_mc_storage, "fig_03b_storage_mc_intervals.png", 10, 7)
```


## Figure 4. EF Distribution Comparison

This summary highlights that storage EFs are generally larger than housing EFs,
while also reminding that the basis differs: TAN for storage and Total N for
housing.

The medium of EF could be used as Tier 1 EF?

```{r}
#| label: fig-ef-distribution
#| fig-width: 10
#| fig-height: 6

housing_distribution <- housing_long %>%
  transmute(
    Stage_Group = "Housing",
    EF = EF_Deterministic
  )

storage_distribution <- storage_plot %>%
  transmute(
    Stage_Group = "Storage",
    EF = Year3_EF_TAN
  )

distribution_data <- bind_rows(housing_distribution, storage_distribution) %>%
  mutate(
    Stage_Group = factor(
      Stage_Group,
      levels = c("Housing", "Storage")
    )
  )


quantile_labels <- distribution_data %>%
  group_by(Stage_Group) %>%
  summarise(
    q25 = quantile(EF, 0.25, na.rm = TRUE),
    median = median(EF, na.rm = TRUE),
    q75 = quantile(EF, 0.75, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  tidyr::pivot_longer(
    cols = c(q25, median, q75),
    names_to = "Statistic",
    values_to = "EF"
  ) %>%
  mutate(
    label = case_when(
      Statistic == "q25" ~ paste0(round(EF, 3)),
      Statistic == "median" ~ paste0(round(EF, 3)),
      Statistic == "q75" ~ paste0(round(EF, 3))
    )
  )


fig_distribution <- ggplot(distribution_data, aes(x = Stage_Group, y = EF)) +
  geom_boxplot(fill = "#d8ead2", color = "#2e4b3d", outlier.alpha = 0.5) +
  geom_jitter(width = 0.13, height = 0, alpha = 0.35, size = 1.1, color = "#a65b3b") +
  geom_text(
    data = quantile_labels,
    aes(x = Stage_Group, y = EF, label = label),
    inherit.aes = FALSE,
    nudge_x = 0.38,
    hjust = 0,
    size = 2.6,
    color = "#2e4b3d"
  ) +
  labs(
    title = "Distribution of deterministic Tier 2 emission factors",
    subtitle = "Housing is Total-N basis; storage is TAN basis",
    x = NULL,
    y = "EF, kg NH3-N / kg TAN or Total N"
  ) +
  theme_tier2(base_size = 9) +
  theme(
    axis.text.x = element_text(angle = 30, hjust = 1),
    plot.margin = margin(5.5, 35, 5.5, 5.5)
  ) +
  coord_cartesian(clip = "off")


fig_distribution
save_plot(fig_distribution, "fig_04_ef_distribution_comparison.png", 10, 6)
```

