---
title: "Proposed Tier 2 Housing NH3 Emission Factors"
format: html
  embed-resources: true
---

## Overview

This file calculates proposed Tier 2 housing ammonia emission factors by
animal type, housing type, manure type, and broad climate zone.

The deterministic output uses broad climate-zone outdoor annual temperature,
converted to an animal-specific indoor housing temperature, and the updated
housing correction-factor table:

```text
EF = EFs_TotalN * CT_15^(T_indoor - T_air)
```

`EFs_TotalN` is treated as the reference EF at the row-specific indoor air
temperature `T_air` from `Inputs/NH3_Correction_Factors.xlsx`.

## Preparation

```{r}
#| label: preparation
#| message: false
#| warning: false

library(readxl)
library(tidyverse)

# Month order used in the climate workbook.
month_labels <- c(
  "Jan", "Feb", "Mar", "Apr", "May", "Jun",
  "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"
)

# Upper cap for emission factors.
# Revise this value if a different maximum EF is needed.
ef_max <- 1

# Monte Carlo settings.
set.seed(2026)
n_iter <- 5000

# Project file paths.
paths <- list(
  factors_xlsx = file.path("Inputs", "NH3_Correction_Factors.xlsx"),
  climate_xlsx = file.path("Inputs", "Climate_zone_temperatures.xlsx"),
  proposed_csv = file.path("Outputs", "Proposed_Tier2_Housing.csv"),
  detail_csv = file.path("Outputs", "Proposed_Tier2_Housing_Detail.csv"),
  mc_summary_csv = file.path("Outputs", "Proposed_Tier2_Housing_MC_Summary.csv")
)

# The final proposal uses four broad climate zones. Moist and dry component
# zones are averaged to represent warm temperate, cool temperate, and boreal.
broad_climate_map <- list(
  Tropical = c("Tropical Wet", "Tropical Moist", "Tropical Dry"),
  Warm_Temperate = c("Warm Temperate Moist", "Warm Temperate Dry"),
  Cool_Temperate = c("Cool Temperate Moist", "Cool Temperate Dry"),
  Boreal = c("Boreal Moist", "Boreal Dry")
)

id_cols <- c("Animal_Type", "Housing", "Manure_type")
climate_cols <- names(broad_climate_map)
```

## Input Data

```{r}
#| label: input-data
#| message: false
#| warning: false

# Read the updated housing EF, reference-temperature, and CT table.
factors <- as.data.frame(
  readxl::read_excel(paths$factors_xlsx, sheet = "Housing"),
  stringsAsFactors = FALSE,
  check.names = FALSE
)

names(factors) <- trimws(names(factors))

required_factor_cols <- c(
  id_cols,
  "EFs_TotalN", "EFs_Lower", "EFs_Upper",
  "T_air", "T_air_Lower", "T_air_Upper",
  "EF_Ref_15", "CT_15", "CT_Lower", "CT_Upper",
  "Note"
)

missing_factor_cols <- setdiff(required_factor_cols, names(factors))
if (length(missing_factor_cols) > 0) {
  stop("Missing Housing-sheet columns: ", paste(missing_factor_cols, collapse = ", "))
}

# Keep the text columns stable for matching and output.
text_cols <- c(id_cols, "Note")
for (col in intersect(text_cols, names(factors))) {
  factors[[col]] <- trimws(as.character(factors[[col]]))
}
factors$Note[is.na(factors$Note) | factors$Note == "NA"] <- ""

# Convert numeric columns from Excel to numeric R vectors.
numeric_cols <- setdiff(required_factor_cols, c(id_cols, "Note"))
for (col in numeric_cols) {
  factors[[col]] <- suppressWarnings(as.numeric(factors[[col]]))
}

# Recreate the output skeleton directly from the updated input sheet.
proposed_result <- factors[id_cols]
```

## Climate Temperatures

```{r}
#| label: climate-temperatures
#| message: false
#| warning: false

# Sheet2 has two header rows: month names in row 1 and Median/Std in row 2.
climate_raw <- readxl::read_excel(
  paths$climate_xlsx,
  sheet = "Sheet2",
  col_names = FALSE,
  .name_repair = "minimal"
)

month_median_cols <- 2 + (seq_along(month_labels) - 1) * 2
month_std_cols <- 3 + (seq_along(month_labels) - 1) * 2

climate_medians <- as.data.frame(
  climate_raw[-c(1, 2), c(1, month_median_cols)],
  stringsAsFactors = FALSE,
  check.names = FALSE
)
climate_stds <- as.data.frame(
  climate_raw[-c(1, 2), c(1, month_std_cols)],
  stringsAsFactors = FALSE,
  check.names = FALSE
)

names(climate_medians) <- c("Climate_zone", month_labels)
names(climate_stds) <- c("Climate_zone", month_labels)

clean_climate_table <- function(climate_table) {
  climate_table$Climate_zone <- trimws(as.character(climate_table$Climate_zone))
  climate_table <- climate_table[
    !is.na(climate_table$Climate_zone) &
      nzchar(climate_table$Climate_zone) &
      climate_table$Climate_zone != "0",
    ,
    drop = FALSE
  ]

  for (month in month_labels) {
    climate_table[[month]] <- suppressWarnings(as.numeric(climate_table[[month]]))
  }

  climate_table
}

climate_medians <- clean_climate_table(climate_medians)
climate_stds <- clean_climate_table(climate_stds)

get_climate_matrix <- function(climate_table, component_zones) {
  climate_rows <- climate_table[climate_table$Climate_zone %in% component_zones, ]

  if (nrow(climate_rows) != length(component_zones)) {
    stop(
      "Expected climate rows for: ",
      paste(component_zones, collapse = ", ")
    )
  }

  as.matrix(climate_rows[month_labels])
}

# Annual outdoor temperatures are calculated after aggregating component climate
# zones into the four broad zones.
broad_climate_summary <- data.frame(
  Climate_zone = climate_cols,
  Outdoor_Annual_Temperature_C = NA_real_,
  Outdoor_Annual_Temperature_SD_C = NA_real_,
  stringsAsFactors = FALSE
)

for (climate_col in climate_cols) {
  component_zones <- broad_climate_map[[climate_col]]
  median_matrix <- get_climate_matrix(climate_medians, component_zones)
  std_matrix <- get_climate_matrix(climate_stds, component_zones)

  monthly_broad_medians <- colMeans(median_matrix, na.rm = TRUE)

  broad_climate_summary[
    broad_climate_summary$Climate_zone == climate_col,
    "Outdoor_Annual_Temperature_C"
  ] <- mean(monthly_broad_medians, na.rm = TRUE)

  # The Monte Carlo uses one annual outdoor temperature draw. Its SD is the
  # simple average of all component-zone monthly Std values for the broad zone.
  broad_climate_summary[
    broad_climate_summary$Climate_zone == climate_col,
    "Outdoor_Annual_Temperature_SD_C"
  ] <- mean(as.numeric(std_matrix), na.rm = TRUE)
}

broad_climate_summary
```

## Calculation Functions

```{r}
#| label: calculation-functions
#| message: false
#| warning: false

cap_ef <- function(ef_value) {
  pmin(ef_value, ef_max)
}

housing_temperature_group <- function(animal_type) {
  animal_type <- tolower(as.character(animal_type))

  if (grepl("cattle", animal_type)) {
    return("cattle")
  }
  if (grepl("swine", animal_type)) {
    return("swine")
  }
  if (grepl("broiler", animal_type)) {
    return("broiler")
  }
  if (grepl("layer", animal_type)) {
    return("layer")
  }

  stop("No indoor temperature equation defined for animal type: ", animal_type)
}

calculate_indoor_temperature <- function(outdoor_temperature_c, animal_type) {
  group <- housing_temperature_group(animal_type)

  if (identical(group, "cattle")) {
    return(outdoor_temperature_c + 3)
  }

  if (identical(group, "swine")) {
    return(ifelse(
      outdoor_temperature_c <= 0,
      20 + 0.5 * outdoor_temperature_c,
      ifelse(
        outdoor_temperature_c <= 12.5,
        20,
        20 + (outdoor_temperature_c - 12.5)
      )
    ))
  }

  if (identical(group, "broiler")) {
    return(
      2.0e-4 * outdoor_temperature_c^3 +
        1.0e-3 * outdoor_temperature_c^2 +
        2.4e-2 * outdoor_temperature_c +
        22.1
    )
  }

  1.4e-4 * outdoor_temperature_c^3 +
    2.3e-3 * outdoor_temperature_c^2 +
    1.1e-2 * outdoor_temperature_c +
    23.8
}

# Notes such as "Only applicable to T >15" define the minimum corrected
# housing temperature required before CT is applied.
threshold_from_note <- function(note) {
  note <- as.character(note)
  match_text <- regmatches(
    note,
    regexpr(">[[:space:]]*[-+]?[0-9]*\\.?[0-9]+", note)
  )

  if (length(match_text) == 0 || is.na(match_text) || match_text == "") {
    return(NA_real_)
  }

  as.numeric(gsub(">|[[:space:]]", "", match_text))
}

ct_is_applicable <- function(ct, corrected_temperature_c, note) {
  if (is.na(ct)) {
    return(FALSE)
  }

  threshold_c <- threshold_from_note(note)
  if (is.na(threshold_c)) {
    return(TRUE)
  }

  corrected_temperature_c > threshold_c
}

calculate_annual_ef <- function(
  ef_ref,
  ct,
  corrected_temperature_c,
  t_air_reference_c,
  note
) {
  if (ct_is_applicable(ct, corrected_temperature_c, note)) {
    corrected_ef <- ef_ref * ct^(corrected_temperature_c - t_air_reference_c)
  } else {
    corrected_ef <- ef_ref
  }

  cap_ef(corrected_ef)
}
```

## Annual EF Calculation

```{r}
#| label: annual-ef-calculation
#| message: false
#| warning: false

deterministic_detail_rows <- list()
deterministic_detail_index <- 1

for (row_index in seq_len(nrow(factors))) {
  for (climate_col in climate_cols) {
    outdoor_annual_temp <- broad_climate_summary[
      broad_climate_summary$Climate_zone == climate_col,
      "Outdoor_Annual_Temperature_C"
    ]
    indoor_annual_temp <- calculate_indoor_temperature(
      outdoor_temperature_c = outdoor_annual_temp,
      animal_type = factors$Animal_Type[row_index]
    )

    ef_value <- calculate_annual_ef(
      ef_ref = factors$EFs_TotalN[row_index],
      ct = factors$CT_15[row_index],
      corrected_temperature_c = indoor_annual_temp,
      t_air_reference_c = factors$T_air[row_index],
      note = factors$Note[row_index]
    )

    proposed_result[row_index, climate_col] <- round(ef_value, 3)

    deterministic_detail_rows[[deterministic_detail_index]] <- data.frame(
      Animal_Type = factors$Animal_Type[row_index],
      Housing = factors$Housing[row_index],
      Manure_type = factors$Manure_type[row_index],
      Climate_zone = climate_col,
      Outdoor_Annual_Temperature_C = round(outdoor_annual_temp, 3),
      Indoor_Annual_Temperature_C = round(indoor_annual_temp, 3),
      T_air_Reference_C = factors$T_air[row_index],
      EFref_TotalN = factors$EFs_TotalN[row_index],
      CT_15 = factors$CT_15[row_index],
      CT_Threshold_C = threshold_from_note(factors$Note[row_index]),
      CT_Applied = ct_is_applicable(
        factors$CT_15[row_index],
        indoor_annual_temp,
        factors$Note[row_index]
      ),
      EF_TotalN = round(ef_value, 3),
      stringsAsFactors = FALSE
    )

    deterministic_detail_index <- deterministic_detail_index + 1
  }
}

deterministic_detail <- do.call(rbind, deterministic_detail_rows)
```

## Monte Carlo Preparation

```{r}
#| label: monte-carlo-preparation
#| message: false
#| warning: false

# Bounded lognormal sampling uses the reported lower and upper 95% CI limits
# as bounds. Values outside the bounds are resampled.
sample_bounded_lognormal <- function(n, center, lower, upper) {
  if (is.na(center) || is.na(lower) || is.na(upper) ||
      center <= 0 || lower <= 0 || upper <= 0 || lower >= upper) {
    return(rep(center, n))
  }

  sdlog <- (log(upper) - log(lower)) / (2 * 1.96)
  samples <- rep(NA_real_, n)
  remaining <- seq_len(n)
  attempts <- 0

  while (length(remaining) > 0 && attempts < 100) {
    draws <- rlnorm(length(remaining), meanlog = log(center), sdlog = sdlog)
    keep <- draws >= lower & draws <= upper

    if (any(keep)) {
      samples[remaining[keep]] <- draws[keep]
    }

    remaining <- remaining[!keep]
    attempts <- attempts + 1
  }

  # If rejection sampling leaves a few empty values, use the point estimate.
  samples[is.na(samples)] <- center
  samples
}

sample_truncated_normal <- function(n, mean, lower, upper) {
  if (is.na(mean) || is.na(lower) || is.na(upper) || lower >= upper) {
    return(rep(mean, n))
  }

  sd <- (upper - lower) / (2 * 1.96)
  if (is.na(sd) || sd <= 0) {
    return(rep(mean, n))
  }

  samples <- rep(NA_real_, n)
  remaining <- seq_len(n)
  attempts <- 0

  while (length(remaining) > 0 && attempts < 100) {
    draws <- rnorm(length(remaining), mean = mean, sd = sd)
    keep <- draws >= lower & draws <= upper

    if (any(keep)) {
      samples[remaining[keep]] <- draws[keep]
    }

    remaining <- remaining[!keep]
    attempts <- attempts + 1
  }

  samples[is.na(samples)] <- mean
  samples
}

outdoor_temperature_samples <- list()
for (climate_col in climate_cols) {
  annual_temp <- broad_climate_summary[
    broad_climate_summary$Climate_zone == climate_col,
    "Outdoor_Annual_Temperature_C"
  ]
  annual_sd <- broad_climate_summary[
    broad_climate_summary$Climate_zone == climate_col,
    "Outdoor_Annual_Temperature_SD_C"
  ]

  outdoor_temperature_samples[[climate_col]] <- rnorm(
    n_iter,
    mean = annual_temp,
    sd = annual_sd
  )
}
```

## Monte Carlo Simulation

```{r}
#| label: monte-carlo-simulation
#| message: false
#| warning: false

mc_summary_rows <- list()
mc_summary_index <- 1

for (row_index in seq_len(nrow(factors))) {
  ef_samples <- sample_bounded_lognormal(
    n = n_iter,
    center = factors$EFs_TotalN[row_index],
    lower = factors$EFs_Lower[row_index],
    upper = factors$EFs_Upper[row_index]
  )

  ct_samples <- sample_bounded_lognormal(
    n = n_iter,
    center = factors$CT_15[row_index],
    lower = factors$CT_Lower[row_index],
    upper = factors$CT_Upper[row_index]
  )

  t_air_samples <- sample_truncated_normal(
    n = n_iter,
    mean = factors$T_air[row_index],
    lower = factors$T_air_Lower[row_index],
    upper = factors$T_air_Upper[row_index]
  )

  threshold_c <- threshold_from_note(factors$Note[row_index])

  for (climate_col in climate_cols) {
    outdoor_temp_samples <- outdoor_temperature_samples[[climate_col]]
    indoor_temp_samples <- calculate_indoor_temperature(
      outdoor_temperature_c = outdoor_temp_samples,
      animal_type = factors$Animal_Type[row_index]
    )

    if (is.na(threshold_c)) {
      ct_applied <- !is.na(ct_samples)
    } else {
      ct_applied <- !is.na(ct_samples) & indoor_temp_samples > threshold_c
    }

    ef_sim <- ifelse(
      ct_applied,
      ef_samples * ct_samples^(indoor_temp_samples - t_air_samples),
      ef_samples
    )
    ef_sim <- cap_ef(ef_sim)

    deterministic_ef <- proposed_result[row_index, climate_col]

    mc_summary_rows[[mc_summary_index]] <- data.frame(
      Animal_Type = factors$Animal_Type[row_index],
      Housing = factors$Housing[row_index],
      Manure_type = factors$Manure_type[row_index],
      Climate_zone = climate_col,
      Outdoor_Annual_Temperature_C = round(
        broad_climate_summary[
          broad_climate_summary$Climate_zone == climate_col,
          "Outdoor_Annual_Temperature_C"
        ],
        3
      ),
      Outdoor_Annual_Temperature_SD_C = round(
        broad_climate_summary[
          broad_climate_summary$Climate_zone == climate_col,
          "Outdoor_Annual_Temperature_SD_C"
        ],
        3
      ),
      Indoor_Annual_Temperature_C = round(
        calculate_indoor_temperature(
          broad_climate_summary[
            broad_climate_summary$Climate_zone == climate_col,
            "Outdoor_Annual_Temperature_C"
          ],
          factors$Animal_Type[row_index]
        ),
        3
      ),
      T_air_Reference_C = factors$T_air[row_index],
      T_air_Lower_C = factors$T_air_Lower[row_index],
      T_air_Upper_C = factors$T_air_Upper[row_index],
      EF_Deterministic = round(deterministic_ef, 3),
      EF_MC_Mean = round(mean(ef_sim, na.rm = TRUE), 3),
      EF_MC_Lower95 = round(as.numeric(quantile(ef_sim, 0.025, na.rm = TRUE)), 3),
      EF_MC_Upper95 = round(as.numeric(quantile(ef_sim, 0.975, na.rm = TRUE)), 3),
      
      CT_Threshold_C = threshold_c,
      stringsAsFactors = FALSE
    )

    mc_summary_index <- mc_summary_index + 1
  }
}

mc_summary <- do.call(rbind, mc_summary_rows)
```

## Monte Carlo Summary 

Expand the summary table for Table 4.3 Tier 2 Housing Emission Factors by animal category, manure management system, climate zone (kg NH3-N kg N-1)

```{r}
#| label: monte-carlo-summary
#| message: false
#| warning: false

head(mc_summary)

mc_summary <- mc_summary %>%
  mutate(`CI_lower%` = round((EF_MC_Lower95 - EF_Deterministic) / EF_Deterministic,2) ,
         `CI_upper%` = round((EF_MC_Upper95 - EF_Deterministic) / EF_Deterministic,2))
          
   

```

## Validation

```{r}
#| label: validation
#| message: false
#| warning: false

expected_output_cols <- c(id_cols, climate_cols)

if (!identical(names(proposed_result), expected_output_cols)) {
  stop("Proposed output columns changed unexpectedly.")
}

if (nrow(proposed_result) != nrow(factors)) {
  stop("Proposed output row count does not match the Housing sheet.")
}

if (nrow(mc_summary) != nrow(factors) * length(climate_cols)) {
  stop("Monte Carlo summary row count changed unexpectedly.")
}

if (!identical(names(proposed_result)[-(seq_along(id_cols))], climate_cols)) {
  stop("Proposed climate-zone columns are not in the expected order.")
}

if (!setequal(unique(mc_summary$Climate_zone), climate_cols)) {
  stop("Monte Carlo summary should contain exactly the four broad climate zones.")
}

if (any(is.na(proposed_result[climate_cols]))) {
  stop("Proposed climate-zone EF columns should not contain missing values.")
}

if (any(is.na(factors$T_air))) {
  stop("Housing T_air reference temperatures should not contain missing values.")
}

if (max(as.matrix(proposed_result[climate_cols]), na.rm = TRUE) > ef_max) {
  stop("Proposed output exceeds ef_max.")
}

mc_ef_cols <- c("EF_Deterministic", "EF_MC_Mean", "EF_MC_Lower95", "EF_MC_Upper95")
if (max(as.matrix(mc_summary[mc_ef_cols]), na.rm = TRUE) > ef_max) {
  stop("Monte Carlo summary exceeds ef_max.")
}
```

## Write Outputs

```{r}
#| label: write-outputs
#| message: false
#| warning: false

write.csv(
  proposed_result,
  paths$proposed_csv,
  row.names = FALSE,
  quote = TRUE,
  na = ""
)

write.csv(
  deterministic_detail,
  paths$detail_csv,
  row.names = FALSE,
  quote = TRUE,
  na = ""
)

write.csv(
  mc_summary,
  paths$mc_summary_csv,
  row.names = FALSE,
  quote = TRUE,
  na = ""
)
```

## Summary

```{r}
#| label: summary
#| message: false
#| warning: false

output_summary <- data.frame(
  output_file = c(paths$proposed_csv, paths$detail_csv, paths$mc_summary_csv),
  rows = c(nrow(proposed_result), nrow(deterministic_detail), nrow(mc_summary)),
  columns = c(ncol(proposed_result), ncol(deterministic_detail), ncol(mc_summary)),
  stringsAsFactors = FALSE
)

output_summary
```

## Method Notes

The final proposal uses annual outdoor temperature only for housing climate
assignment. Monthly temperature medians are used only to construct annual broad
outdoor temperatures from the climate workbook.

Broad climate zones are represented by averaging component climate-zone monthly
medians before calculating the annual mean. Warm temperate and cool temperate
therefore average their moist and dry component zones.

For the Monte Carlo simulation, monthly outdoor temperatures are not sampled
independently. Instead, one annual outdoor temperature is sampled for each broad
climate zone using the annual mean temperature and an annual-level standard
deviation derived from the workbook's monthly `Std` values. The SD is treated as
a representative broad-zone variability term, not as a formal uncertainty of the
mean climate-zone average. Each outdoor draw is then converted to animal-specific
indoor temperature before EF correction.

The correction uses `EFs_TotalN` as the row-specific `T_air` reference EF.
Monte Carlo simulations also sample `T_air` from `T_air_Lower` and
`T_air_Upper`. The input column `EF_Ref_15` is retained in the input table but
is not used in the main proposed output.
