---
title: "DATAMAN-AMCLM Implied Temperature Check"
format: html
---

## Overview

This file back-calculates the temperature implied by each DATAMAN/global
reference housing EF (`EFs_TotalN`) when compared with the modelled AMCLIM
reference EF at 15 degrees C (`EF_Ref_15`) and the temperature correction
factor (`CT_15`).

The calculation rearranges the exponential equation:

```text
EFs_TotalN = EF_Ref_15 * CT_15^(T_implied - 15)
T_implied = 15 + log(EFs_TotalN / EF_Ref_15) / log(CT_15)
```

This is a diagnostic back-calculation, not a new EF estimate.

## Preparation

```{r}
#| label: preparation
#| message: false
#| warning: false

library(readxl)

# Project file paths.
paths <- list(
  factors_xlsx = file.path("Inputs", "NH3_Correction_Factors.xlsx"),
  output_csv = file.path("Outputs", "DATAMAN_AMCLM_Implied_Temperature.csv")
)

# Temperature flags used to screen implied temperatures.
very_low_temperature_c <- 5
very_high_temperature_c <- 30

id_cols <- c("Animal_Type", "Housing", "Manure_type")
required_cols <- c(id_cols, "EFs_TotalN", "EF_Ref_15", "CT_15", "Note")
```

## Input Data

```{r}
#| label: input-data
#| message: false
#| warning: false

# Read the updated housing EF and CT table.
factors <- as.data.frame(
  readxl::read_excel(paths$factors_xlsx, sheet = "Housing"),
  stringsAsFactors = FALSE,
  check.names = FALSE
)

names(factors) <- trimws(names(factors))

missing_cols <- setdiff(required_cols, names(factors))
if (length(missing_cols) > 0) {
  stop("Missing required input columns: ", paste(missing_cols, collapse = ", "))
}

# Keep text values stable for output.
for (col in c(id_cols, "Note")) {
  factors[[col]] <- trimws(as.character(factors[[col]]))
}
factors$Note[is.na(factors$Note) | factors$Note == "NA"] <- ""

# Convert the values needed by the inverse equation to numeric vectors.
for (col in c("EFs_TotalN", "EF_Ref_15", "CT_15")) {
  factors[[col]] <- suppressWarnings(as.numeric(factors[[col]]))
}
```

## Calculation

```{r}
#| label: calculation
#| message: false
#| warning: false

# Notes such as "Only applicable to T >15" define the minimum temperature
# required before the CT relationship is expected to apply.
threshold_from_note <- function(note) {
  note <- as.character(note)
  match_text <- regmatches(
    note,
    regexpr(">[[:space:]]*[-+]?[0-9]*\\.?[0-9]+", note)
  )

  if (length(match_text) == 0 || is.na(match_text) || match_text == "") {
    return(NA_real_)
  }

  as.numeric(gsub(">|[[:space:]]", "", match_text))
}

can_calculate_implied_temperature <- function(ef_total_n, ef_ref_15, ct_15) {
  !is.na(ef_total_n) &&
    !is.na(ef_ref_15) &&
    !is.na(ct_15) &&
    ef_total_n > 0 &&
    ef_ref_15 > 0 &&
    ct_15 > 0 &&
    ct_15 != 1
}

calculate_implied_temperature <- function(ef_total_n, ef_ref_15, ct_15) {
  if (!can_calculate_implied_temperature(ef_total_n, ef_ref_15, ct_15)) {
    return(NA_real_)
  }

  15 + log(ef_total_n / ef_ref_15) / log(ct_15)
}

implied_temperature <- vapply(
  seq_len(nrow(factors)),
  function(row_index) {
    calculate_implied_temperature(
      ef_total_n = factors$EFs_TotalN[row_index],
      ef_ref_15 = factors$EF_Ref_15[row_index],
      ct_15 = factors$CT_15[row_index]
    )
  },
  numeric(1)
)

ct_threshold <- vapply(factors$Note, threshold_from_note, numeric(1))
```

## Threshold Flagging

```{r}
#| label: threshold-flagging
#| message: false
#| warning: false

build_temperature_flag <- function(
  implied_temp,
  threshold_c,
  ef_total_n,
  ef_ref_15,
  ct_15
) {
  if (!can_calculate_implied_temperature(ef_total_n, ef_ref_15, ct_15)) {
    return("Cannot calculate")
  }

  flags <- character(0)

  if (!is.na(threshold_c) && implied_temp <= threshold_c) {
    flags <- c(flags, "Below CT threshold")
  }

  if (implied_temp < very_low_temperature_c) {
    flags <- c(flags, "Very low implied temperature")
  }

  if (implied_temp > very_high_temperature_c) {
    flags <- c(flags, "Very high implied temperature")
  }

  if (length(flags) == 0) {
    return("OK")
  }

  paste(flags, collapse = "; ")
}

implied_temperature_flag <- vapply(
  seq_len(nrow(factors)),
  function(row_index) {
    build_temperature_flag(
      implied_temp = implied_temperature[row_index],
      threshold_c = ct_threshold[row_index],
      ef_total_n = factors$EFs_TotalN[row_index],
      ef_ref_15 = factors$EF_Ref_15[row_index],
      ct_15 = factors$CT_15[row_index]
    )
  },
  character(1)
)

implied_temperature_output <- data.frame(
  Animal_Type = factors$Animal_Type,
  Housing = factors$Housing,
  Manure_type = factors$Manure_type,
  EFs_TotalN = factors$EFs_TotalN,
  EF_Ref_15 = factors$EF_Ref_15,
  CT_15 = factors$CT_15,
  Implied_Temperature_C = round(implied_temperature, 3),
  Implied_Temperature_Flag = implied_temperature_flag,
  Note = factors$Note,
  stringsAsFactors = FALSE
)
```

## Diagnostics

```{r}
#| label: diagnostics
#| message: false
#| warning: false

valid_temperatures <- implied_temperature[is.finite(implied_temperature)]

diagnostic_summary <- data.frame(
  Metric = c(
    "Rows",
    "Minimum implied temperature C",
    "Median implied temperature C",
    "Maximum implied temperature C",
    "Rows below 5 C",
    "Rows above 30 C",
    "Rows below CT applicability threshold"
  ),
  Value = c(
    nrow(implied_temperature_output),
    round(min(valid_temperatures), 3),
    round(median(valid_temperatures), 3),
    round(max(valid_temperatures), 3),
    sum(valid_temperatures < very_low_temperature_c),
    sum(valid_temperatures > very_high_temperature_c),
    sum(!is.na(ct_threshold) & implied_temperature <= ct_threshold, na.rm = TRUE)
  ),
  stringsAsFactors = FALSE
)

diagnostic_summary
```

## Validation

```{r}
#| label: validation
#| message: false
#| warning: false

if (nrow(implied_temperature_output) != nrow(factors)) {
  stop("Output row count does not match the Housing sheet.")
}

if (any(is.na(implied_temperature_output$Implied_Temperature_C) &
        implied_temperature_output$Implied_Temperature_Flag != "Cannot calculate")) {
  stop("Missing implied temperatures must be flagged as Cannot calculate.")
}

if (any(grepl("Below CT threshold", implied_temperature_output$Implied_Temperature_Flag)) &&
    !any(!is.na(ct_threshold))) {
  stop("Threshold flags were assigned even though no thresholds were found.")
}

if (!any(grepl("Very low implied temperature", implied_temperature_output$Implied_Temperature_Flag))) {
  warning("No rows were flagged as very low implied temperature.")
}

if (!any(grepl("Very high implied temperature", implied_temperature_output$Implied_Temperature_Flag))) {
  warning("No rows were flagged as very high implied temperature.")
}
```

## Write Outputs

```{r}
#| label: write-outputs
#| message: false
#| warning: false

dir.create(dirname(paths$output_csv), showWarnings = FALSE, recursive = TRUE)

write.csv(
  implied_temperature_output,
  paths$output_csv,
  row.names = FALSE,
  quote = TRUE,
  na = ""
)
```

## Summary

```{r}
#| label: summary
#| message: false
#| warning: false

head(implied_temperature_output)
diagnostic_summary
```

## Interpretation Note

Very low or very high implied temperatures suggest that `EFs_TotalN`,
`EF_Ref_15`, and `CT_15` may not share the same reference basis for that row.
For rows with CT threshold notes, implied temperatures below the threshold also
suggest that the exponential CT relationship may not be applicable to the
back-calculated value.
