---
title: "Proposed Tier 2 Storage NH3 Emission Factors"
format: html
embed-resources: true
---

## Overview

This file calculates proposed Tier 2 manure storage ammonia emission factors by
animal type, storage type, manure type, and broad climate zone.

Storage uses a TAN mass-balance approach. Each month a fixed amount of TAN
enters storage, temperature-corrected NH3 loss is calculated from available TAN,
and remaining TAN is carried forward. The reported annual EF is calculated from
the third simulation year. Only the April and October removal schedule is used.

## Preparation

```{r}
#| label: preparation
#| message: false
#| warning: false

library(readxl)
library(tidyverse)

# Month order used in the climate workbook and storage simulation.
month_labels <- c(
  "Jan", "Feb", "Mar", "Apr", "May", "Jun",
  "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"
)

month_numbers <- seq_along(month_labels)
names(month_numbers) <- month_labels

# -------------------------------------------------------------------------
# Editable model parameters
# -------------------------------------------------------------------------
seed <- 2026
n_iter <- 5000
monthly_tan_kg <- 100
simulation_years <- 3
removal_efficiency_mean <- 0.95
removal_efficiency_sd <- 0.03
ef_max <- 1

removal_schedules <- list(
  April_October = c(month_numbers["Apr"], month_numbers["Oct"])
)

paths <- list(
  factors_xlsx = file.path("Inputs", "NH3_Correction_Factors.xlsx"),
  climate_xlsx = file.path("Inputs", "Climate_zone_temperatures.xlsx"),
  proposed_csv = file.path("Outputs", "Proposed_Tier2_Storage.csv"),
  mc_summary_csv = file.path("Outputs", "Proposed_Tier2_Storage_MC_Summary.csv"),
  mc_temperatures_csv = file.path(
    "Outputs",
    "Proposed_Tier2_Storage_MC_Temperatures.csv"
  )
)

set.seed(seed)

id_cols <- c("Animal_Type", "Storage", "Manure_type")
total_months <- simulation_years * length(month_labels)
year3_tan_input_kg <- length(month_labels) * monthly_tan_kg

# Storage uses the same four broad climate zones as housing.
broad_climate_map <- list(
  Tropical = c("Tropical Wet", "Tropical Moist", "Tropical Dry"),
  Warm_Temperate = c("Warm Temperate Moist", "Warm Temperate Dry"),
  Cool_Temperate = c("Cool Temperate Moist", "Cool Temperate Dry"),
  Boreal = c("Boreal Moist", "Boreal Dry")
)
climate_zones <- names(broad_climate_map)

dir.create("Outputs", showWarnings = FALSE, recursive = TRUE)
```

## Input Data

```{r}
#| label: input-data
#| message: false
#| warning: false

factors <- as.data.frame(
  readxl::read_excel(paths$factors_xlsx, sheet = "Storage"),
  stringsAsFactors = FALSE,
  check.names = FALSE
)

names(factors) <- trimws(names(factors))

required_factor_cols <- c(
  id_cols,
  "EFs_TAN", "EFs_Lower", "EFs_Upper",
  "T_air", "T_air_Lower", "T_air_Upper",
  "EF_Ref_15", "CT", "CT_Lower", "CT_Upper"
)

missing_factor_cols <- setdiff(required_factor_cols, names(factors))
if (length(missing_factor_cols) > 0) {
  stop("Missing Storage-sheet columns: ", paste(missing_factor_cols, collapse = ", "))
}

for (col in id_cols) {
  factors[[col]] <- trimws(as.character(factors[[col]]))
}

numeric_factor_cols <- setdiff(required_factor_cols, id_cols)
for (col in numeric_factor_cols) {
  factors[[col]] <- suppressWarnings(as.numeric(factors[[col]]))
}

# Sheet2 has two header rows: month names in row 1 and Median/Std in row 2.
climate_raw <- readxl::read_excel(
  paths$climate_xlsx,
  sheet = "Sheet2",
  col_names = FALSE,
  .name_repair = "minimal"
)

month_median_cols <- 2 + (seq_along(month_labels) - 1) * 2
month_std_cols <- 3 + (seq_along(month_labels) - 1) * 2

climate_medians <- as.data.frame(
  climate_raw[-c(1, 2), c(1, month_median_cols)],
  stringsAsFactors = FALSE,
  check.names = FALSE
)
climate_stds <- as.data.frame(
  climate_raw[-c(1, 2), c(1, month_std_cols)],
  stringsAsFactors = FALSE,
  check.names = FALSE
)

names(climate_medians) <- c("Climate_zone", month_labels)
names(climate_stds) <- c("Climate_zone", month_labels)

clean_climate_table <- function(climate_table) {
  climate_table$Climate_zone <- trimws(as.character(climate_table$Climate_zone))
  climate_table <- climate_table[
    !is.na(climate_table$Climate_zone) &
      nzchar(climate_table$Climate_zone) &
      climate_table$Climate_zone != "0",
    ,
    drop = FALSE
  ]

  for (month in month_labels) {
    climate_table[[month]] <- suppressWarnings(as.numeric(climate_table[[month]]))
  }

  row.names(climate_table) <- NULL
  climate_table
}

climate_medians <- clean_climate_table(climate_medians)
climate_stds <- clean_climate_table(climate_stds)

get_climate_matrix <- function(climate_table, component_zones) {
  climate_rows <- climate_table[climate_table$Climate_zone %in% component_zones, ]

  if (nrow(climate_rows) != length(component_zones)) {
    stop(
      "Expected climate rows for: ",
      paste(component_zones, collapse = ", ")
    )
  }

  as.matrix(climate_rows[month_labels])
}

broad_climate_medians <- data.frame(
  Climate_zone = climate_zones,
  matrix(NA_real_, nrow = length(climate_zones), ncol = length(month_labels)),
  stringsAsFactors = FALSE,
  check.names = FALSE
)
broad_climate_stds <- broad_climate_medians
names(broad_climate_medians) <- c("Climate_zone", month_labels)
names(broad_climate_stds) <- c("Climate_zone", month_labels)

for (climate_zone in climate_zones) {
  component_zones <- broad_climate_map[[climate_zone]]
  median_matrix <- get_climate_matrix(climate_medians, component_zones)
  std_matrix <- get_climate_matrix(climate_stds, component_zones)

  broad_climate_medians[
    broad_climate_medians$Climate_zone == climate_zone,
    month_labels
  ] <- as.list(colMeans(median_matrix, na.rm = TRUE))

  broad_climate_stds[
    broad_climate_stds$Climate_zone == climate_zone,
    month_labels
  ] <- as.list(colMeans(std_matrix, na.rm = TRUE))
}

factors
broad_climate_medians
```

## Calculation Functions

```{r}
#| label: calculation-functions
#| message: false
#| warning: false

cap_ef <- function(ef_value) {
  pmin(ef_value, ef_max)
}

calculate_monthly_ef <- function(ef_ref, ct, temperatures_c, t_air_reference_c) {
  if (is.na(ct)) {
    return(cap_ef(rep(ef_ref, length(temperatures_c))))
  }

  cap_ef(ef_ref * ct^(temperatures_c - t_air_reference_c))
}

simulate_storage_annual_ef <- function(
  ef_matrix,
  removal_months,
  removal_efficiency
) {
  if (!is.matrix(ef_matrix)) {
    ef_matrix <- matrix(ef_matrix, nrow = 1)
  }

  n_sim <- nrow(ef_matrix)

  if (length(removal_efficiency) == 1) {
    removal_efficiency <- rep(removal_efficiency, n_sim)
  }

  stored_tan <- rep(0, n_sim)
  year3_nh3 <- rep(0, n_sim)
  year3_tan <- rep(0, n_sim)
  year3_weighted_ef_sum <- rep(0, n_sim)
  year3_available_tan_sum <- rep(0, n_sim)
  year3_unweighted_monthly_ef_sum <- rep(0, n_sim)
  year3_month_count <- 0
  minimum_stored_tan <- rep(Inf, n_sim)

  for (period_index in seq_len(total_months)) {
    month_index <- ((period_index - 1) %% length(month_labels)) + 1
    year_index <- ((period_index - 1) %/% length(month_labels)) + 1

    if (month_index %in% removal_months) {
      # removal_efficiency is the fraction removed from storage; the balance remains.
      stored_tan <- stored_tan * (1 - removal_efficiency)
    }

    available_tan <- stored_tan + monthly_tan_kg
    nh3_loss <- pmin(available_tan, available_tan * ef_matrix[, period_index])
    monthly_storage_ef <- ifelse(
      available_tan > 0,
      nh3_loss / available_tan,
      NA_real_
    )
    stored_tan <- available_tan - nh3_loss
    minimum_stored_tan <- pmin(minimum_stored_tan, stored_tan)

    if (year_index == simulation_years) {
      year3_nh3 <- year3_nh3 + nh3_loss
      year3_tan <- year3_tan + monthly_tan_kg
      year3_weighted_ef_sum <- year3_weighted_ef_sum +
        monthly_storage_ef * available_tan
      year3_available_tan_sum <- year3_available_tan_sum + available_tan
      year3_unweighted_monthly_ef_sum <- year3_unweighted_monthly_ef_sum +
        monthly_storage_ef
      year3_month_count <- year3_month_count + 1
    }
  }

  list(
    annual_ef = year3_weighted_ef_sum / year3_available_tan_sum,
    year3_unweighted_monthly_ef =
      year3_unweighted_monthly_ef_sum / year3_month_count,
    year3_mass_balance_ef = year3_nh3 / year3_tan,
    year3_nh3 = year3_nh3,
    year3_tan = year3_tan,
    year3_available_tan = year3_available_tan_sum,
    final_stored_tan = stored_tan,
    minimum_stored_tan = minimum_stored_tan
  )
}

sample_bounded_lognormal <- function(n, center, lower, upper) {
  if (is.na(center) || is.na(lower) || is.na(upper) ||
      center <= 0 || lower <= 0 || upper <= 0 || lower >= upper) {
    return(rep(center, n))
  }

  sdlog <- (log(upper) - log(lower)) / (2 * 1.96)
  samples <- rep(NA_real_, n)
  remaining <- seq_len(n)
  attempts <- 0

  while (length(remaining) > 0 && attempts < 100) {
    draws <- rlnorm(length(remaining), meanlog = log(center), sdlog = sdlog)
    keep <- draws >= lower & draws <= upper

    if (any(keep)) {
      samples[remaining[keep]] <- draws[keep]
    }

    remaining <- remaining[!keep]
    attempts <- attempts + 1
  }

  samples[is.na(samples)] <- center
  samples
}

sample_truncated_normal <- function(n, mean, sd, lower = 0, upper = 1) {
  if (is.na(sd) || sd <= 0) {
    return(rep(min(max(mean, lower), upper), n))
  }

  samples <- rep(NA_real_, n)
  remaining <- seq_len(n)
  attempts <- 0

  while (length(remaining) > 0 && attempts < 100) {
    draws <- rnorm(length(remaining), mean = mean, sd = sd)
    keep <- draws >= lower & draws <= upper

    if (any(keep)) {
      samples[remaining[keep]] <- draws[keep]
    }

    remaining <- remaining[!keep]
    attempts <- attempts + 1
  }

  samples[is.na(samples)] <- min(max(mean, lower), upper)
  samples
}
```

## Deterministic Storage EF

```{r}
#| label: deterministic-storage-ef
#| message: false
#| warning: false

deterministic_rows <- list()
deterministic_index <- 1

for (row_index in seq_len(nrow(factors))) {
  for (schedule_name in names(removal_schedules)) {
    removal_months <- removal_schedules[[schedule_name]]
    removal_month_names <- paste(month_labels[removal_months], collapse = ";")

    for (climate_zone in climate_zones) {
      temperatures <- as.numeric(
        broad_climate_medians[
          broad_climate_medians$Climate_zone == climate_zone,
          month_labels
        ]
      )

      monthly_ef <- calculate_monthly_ef(
        ef_ref = factors$EFs_TAN[row_index],
        ct = factors$CT[row_index],
        temperatures_c = temperatures,
        t_air_reference_c = factors$T_air[row_index]
      )

      ef_matrix <- matrix(rep(monthly_ef, simulation_years), nrow = 1)

      simulated <- simulate_storage_annual_ef(
        ef_matrix = ef_matrix,
        removal_months = removal_months,
        removal_efficiency = removal_efficiency_mean
      )

      deterministic_rows[[deterministic_index]] <- data.frame(
        Animal_Type = factors$Animal_Type[row_index],
        Storage = factors$Storage[row_index],
        Manure_type = factors$Manure_type[row_index],
        Removal_Schedule = schedule_name,
        Removal_Months = removal_month_names,
        Climate_zone = climate_zone,
        CT = factors$CT[row_index],
        CT_Applied = !is.na(factors$CT[row_index]),
        EFs_TAN = factors$EFs_TAN[row_index],
        T_air_Reference_C = factors$T_air[row_index],
        Annual_Outdoor_Temperature_C = round(mean(temperatures, na.rm = TRUE), 3),
        Monthly_TAN_kg = monthly_tan_kg,
        Removal_Efficiency = removal_efficiency_mean,
        Year3_TAN_Input_kg = simulated$year3_tan,
        Year3_TAN_Available_kg = simulated$year3_available_tan,
        Year3_NH3_N_kg = simulated$year3_nh3,
        Year3_EF_TAN = round(simulated$annual_ef, 3),
        Year3_Unweighted_Monthly_EF_TAN = round(
          simulated$year3_unweighted_monthly_ef,
          3
        ),
        Year3_Mass_Balance_EF_TAN = round(simulated$year3_mass_balance_ef, 3),
        Final_Stored_TAN_kg = simulated$final_stored_tan,
        stringsAsFactors = FALSE
      )

      deterministic_index <- deterministic_index + 1
    }
  }
}

proposed_result <- do.call(rbind, deterministic_rows)

head(proposed_result)
```

## Monte Carlo Temperature Samples

```{r}
#| label: monte-carlo-temperature-samples
#| message: false
#| warning: false

temperature_sample_rows <- vector(
  "list",
  length(climate_zones)
)

for (climate_index in seq_along(climate_zones)) {
  climate_zone <- climate_zones[climate_index]
  n_rows <- n_iter * simulation_years

  climate_samples <- data.frame(
    Iteration = rep(seq_len(n_iter), each = simulation_years),
    Climate_zone = rep(climate_zone, n_rows),
    Year = rep(seq_len(simulation_years), times = n_iter),
    stringsAsFactors = FALSE
  )

  for (month in month_labels) {
    month_median <- broad_climate_medians[
      broad_climate_medians$Climate_zone == climate_zone,
      month
    ]
    month_std <- broad_climate_stds[
      broad_climate_stds$Climate_zone == climate_zone,
      month
    ]

    climate_samples[[month]] <- rnorm(
      n_rows,
      mean = as.numeric(month_median),
      sd = as.numeric(month_std)
    )
  }

  temperature_sample_rows[[climate_index]] <- climate_samples
}

mc_temperatures <- do.call(rbind, temperature_sample_rows)
row.names(mc_temperatures) <- NULL

temperature_matrices <- list()
for (climate_zone in climate_zones) {
  climate_samples <- mc_temperatures[
    mc_temperatures$Climate_zone == climate_zone,
    ,
    drop = FALSE
  ]

  temp_matrix <- matrix(NA_real_, nrow = n_iter, ncol = total_months)

  for (year_index in seq_len(simulation_years)) {
    year_samples <- climate_samples[
      climate_samples$Year == year_index,
      ,
      drop = FALSE
    ]
    year_samples <- year_samples[order(year_samples$Iteration), ]

    month_cols <- ((year_index - 1) * length(month_labels) + 1):
      (year_index * length(month_labels))
    temp_matrix[, month_cols] <- as.matrix(year_samples[month_labels])
  }

  temperature_matrices[[climate_zone]] <- temp_matrix
}

head(mc_temperatures)
```

## Monte Carlo Simulation

```{r}
#| label: monte-carlo-simulation
#| message: false
#| warning: false

removal_efficiency_samples <- sample_truncated_normal(
  n = n_iter,
  mean = removal_efficiency_mean,
  sd = removal_efficiency_sd,
  lower = 0,
  upper = 1
)

mc_summary_rows <- list()
mc_summary_index <- 1

for (row_index in seq_len(nrow(factors))) {
  ef_samples <- sample_bounded_lognormal(
    n = n_iter,
    center = factors$EFs_TAN[row_index],
    lower = factors$EFs_Lower[row_index],
    upper = factors$EFs_Upper[row_index]
  )

  ct_samples <- sample_bounded_lognormal(
    n = n_iter,
    center = factors$CT[row_index],
    lower = factors$CT_Lower[row_index],
    upper = factors$CT_Upper[row_index]
  )

  t_air_samples <- sample_truncated_normal(
    n = n_iter,
    mean = factors$T_air[row_index],
    sd = (factors$T_air_Upper[row_index] - factors$T_air_Lower[row_index]) / (2 * 1.96),
    lower = factors$T_air_Lower[row_index],
    upper = factors$T_air_Upper[row_index]
  )

  for (schedule_name in names(removal_schedules)) {
    removal_months <- removal_schedules[[schedule_name]]
    removal_month_names <- paste(month_labels[removal_months], collapse = ";")

    for (climate_zone in climate_zones) {
      ct_center <- factors$CT[row_index]
      temp_matrix <- temperature_matrices[[climate_zone]]

      if (is.na(ct_center)) {
        ef_matrix <- matrix(ef_samples, nrow = n_iter, ncol = total_months)
      } else {
        ef_matrix <- ef_samples *
          ct_samples^(temp_matrix - t_air_samples)
      }

      ef_matrix <- cap_ef(ef_matrix)

      simulated <- simulate_storage_annual_ef(
        ef_matrix = ef_matrix,
        removal_months = removal_months,
        removal_efficiency = removal_efficiency_samples
      )

      deterministic_ef <- proposed_result[
        proposed_result$Animal_Type == factors$Animal_Type[row_index] &
          proposed_result$Storage == factors$Storage[row_index] &
          proposed_result$Manure_type == factors$Manure_type[row_index] &
          proposed_result$Removal_Schedule == schedule_name &
          proposed_result$Climate_zone == climate_zone,
        "Year3_EF_TAN"
      ]

      mc_summary_rows[[mc_summary_index]] <- data.frame(
        Animal_Type = factors$Animal_Type[row_index],
        Storage = factors$Storage[row_index],
        Manure_type = factors$Manure_type[row_index],
        Removal_Schedule = schedule_name,
        Removal_Months = removal_month_names,
        Climate_zone = climate_zone,
        CT_Applied = !is.na(ct_center),
        CT = ct_center,
        EFs_TAN = factors$EFs_TAN[row_index],
        T_air_Reference_C = factors$T_air[row_index],
        T_air_Lower_C = factors$T_air_Lower[row_index],
        T_air_Upper_C = factors$T_air_Upper[row_index],
        Monthly_TAN_kg = monthly_tan_kg,
        Removal_Efficiency_Mean = removal_efficiency_mean,
        Removal_Efficiency_SD = removal_efficiency_sd,
        Year3_TAN_Input_kg = year3_tan_input_kg,
        TAN_Available_MC_Mean_kg = round(
          mean(simulated$year3_available_tan, na.rm = TRUE),
          3
        ),
        EF_Deterministic = deterministic_ef,
        EF_MC_Mean = round(mean(simulated$annual_ef, na.rm = TRUE), 3),
        EF_MC_Lower95 = round(
          as.numeric(quantile(simulated$annual_ef, 0.025, na.rm = TRUE)),
          3
        ),
        EF_MC_Upper95 = round(
          as.numeric(quantile(simulated$annual_ef, 0.975, na.rm = TRUE)),
          3
        ),
        Unweighted_Monthly_EF_MC_Mean = round(
          mean(simulated$year3_unweighted_monthly_ef, na.rm = TRUE),
          3
        ),
        Mass_Balance_EF_MC_Mean = round(
          mean(simulated$year3_mass_balance_ef, na.rm = TRUE),
          3
        ),
        NH3_N_MC_Mean_kg = round(mean(simulated$year3_nh3, na.rm = TRUE), 3),
        Final_Stored_TAN_MC_Mean_kg = round(
          mean(simulated$final_stored_tan, na.rm = TRUE),
          3
        ),
        stringsAsFactors = FALSE
      )

      mc_summary_index <- mc_summary_index + 1
    }
  }
}

mc_summary <- do.call(rbind, mc_summary_rows)

head(mc_summary)

#Calculate CI% for Table 4.4 Tier 2 Manure storage emission factors by animal category, manure management system
mc_summary <- mc_summary %>%
  mutate(`CI_lower%` = round((EF_MC_Lower95 - EF_Deterministic) / EF_Deterministic,2) ,
         `CI_upper%` = round((EF_MC_Upper95 - EF_Deterministic) / EF_Deterministic,2))

```

## Validation

```{r}
#| label: validation
#| message: false
#| warning: false

expected_result_rows <- nrow(factors) * length(removal_schedules) * length(climate_zones)
expected_temperature_rows <- n_iter * length(climate_zones) * simulation_years

if (nrow(proposed_result) != expected_result_rows) {
  stop("Unexpected deterministic output rows: ", nrow(proposed_result))
}

if (nrow(mc_summary) != expected_result_rows) {
  stop("Unexpected Monte Carlo summary rows: ", nrow(mc_summary))
}

if (nrow(mc_temperatures) != expected_temperature_rows) {
  stop("Unexpected Monte Carlo temperature rows: ", nrow(mc_temperatures))
}

if (!setequal(proposed_result$Climate_zone, climate_zones) ||
    !setequal(mc_summary$Climate_zone, climate_zones) ||
    !setequal(mc_temperatures$Climate_zone, climate_zones)) {
  stop("Storage outputs should contain exactly the four broad climate zones.")
}

if (!identical(unique(proposed_result$Climate_zone), climate_zones) ||
    !identical(unique(mc_summary$Climate_zone), climate_zones) ||
    !identical(unique(mc_temperatures$Climate_zone), climate_zones)) {
  stop("Storage broad climate zones are not in the expected order.")
}

if (any(grepl("CT_wet|CT_dry", names(proposed_result))) ||
    any(grepl("CT_wet|CT_dry", names(mc_summary)))) {
  stop("Storage outputs should not include wet/dry CT columns.")
}

if (any(is.na(factors$T_air))) {
  stop("Storage T_air reference temperatures should not contain missing values.")
}

if (any(!is.finite(proposed_result$Year3_EF_TAN))) {
  stop("Deterministic storage EF contains non-finite values.")
}

if (any(proposed_result$Year3_EF_TAN < 0) ||
    any(proposed_result$Year3_EF_TAN > ef_max)) {
  stop("Deterministic storage EF is outside expected bounds.")
}

mc_ef_cols <- c("EF_MC_Mean", "EF_MC_Lower95", "EF_MC_Upper95")
if (any(!is.finite(as.matrix(mc_summary[mc_ef_cols])))) {
  stop("Monte Carlo summary contains non-finite EF values.")
}

if (any(as.matrix(mc_summary[mc_ef_cols]) < 0) ||
    any(as.matrix(mc_summary[mc_ef_cols]) > ef_max)) {
  stop("Monte Carlo summary EF is outside expected bounds.")
}

if (any(abs(proposed_result$Year3_TAN_Input_kg - year3_tan_input_kg) > 1e-9)) {
  stop("Year 3 TAN denominator changed unexpectedly.")
}

if (any(proposed_result$Final_Stored_TAN_kg < -1e-9)) {
  stop("Negative stored TAN detected in deterministic results.")
}
```

## Write Outputs

```{r}
#| label: write-outputs
#| message: false
#| warning: false

write.csv(
  proposed_result,
  paths$proposed_csv,
  row.names = FALSE,
  quote = TRUE,
  na = ""
)

write.csv(
  mc_summary,
  paths$mc_summary_csv,
  row.names = FALSE,
  quote = TRUE,
  na = ""
)

write.csv(
  mc_temperatures,
  paths$mc_temperatures_csv,
  row.names = FALSE,
  quote = TRUE,
  na = ""
)
```

## Summary

```{r}
#| label: summary
#| message: false
#| warning: false

output_summary <- data.frame(
  output_file = c(
    paths$proposed_csv,
    paths$mc_summary_csv,
    paths$mc_temperatures_csv
  ),
  rows = c(nrow(proposed_result), nrow(mc_summary), nrow(mc_temperatures)),
  columns = c(ncol(proposed_result), ncol(mc_summary), ncol(mc_temperatures)),
  stringsAsFactors = FALSE
)

output_summary
```

## Method Notes

The storage calculation uses `EFs_TAN` as the reference-temperature emission
factor at the row-specific outdoor storage reference temperature `T_air`.
Monthly broad climate-zone outdoor temperatures are corrected with:

```text
EF_month = EFs_TAN * CT^(Temperature_month - T_air)
```

Broad climate zones are represented by averaging component climate-zone monthly
medians. Warm temperate and cool temperate average their moist and dry component
zones, and no wet/dry CT subzone distinction is used. Monthly `Std` values are
averaged across component zones and treated as representative broad-zone
variability, not as a formal uncertainty of the broad-zone mean.

Only the April and October removal schedule is used because prior comparisons
showed storage EFs were not sensitive to the other removal scenarios.

The Monte Carlo temperature output is written in a reusable wide format with one
row per iteration, broad climate zone, and simulation year, and one column for
each month. Monte Carlo simulations also sample the row-specific reference
temperature `T_air` from `T_air_Lower` and `T_air_Upper`.
