---
title: "QC: Proposed Tier 2 Storage Monthly TAN and NH3"
format:
  html:
    toc: true
    toc-depth: 2
---

## Overview

This QC file shows the month-by-month storage mass balance used for the
proposed Tier 2 storage NH3 emission factor calculation.

The QC scenarios are:

- Dairy Cattle, Tank, Slurry
- Dairy Cattle, Manure heap, Solid

Both are evaluated for `Cool Temperate Moist` under all four removal schedules.

## Preparation

```{r}
#| label: preparation
#| message: false
#| warning: false

library(dplyr)
library(ggplot2)
library(knitr)
library(readxl)
library(readr)

month_labels <- c(
  "Jan", "Feb", "Mar", "Apr", "May", "Jun",
  "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"
)

month_numbers <- seq_along(month_labels)
names(month_numbers) <- month_labels

# -------------------------------------------------------------------------
# Editable QC parameters
# -------------------------------------------------------------------------
monthly_tan_kg <- 100
simulation_years <- 3
reference_temperature_c <- 15
removal_efficiency <- 0.95
ef_max <- 1
qc_climate_zone <- "Cool Temperate Dry"

qc_combinations <- data.frame(
  Animal_Type = c("Dairy Cattle", "Dairy Cattle"),
  Storage = c("Tank", "Manure heap"),
  Manure_type = c("Slurry", "Solid"),
  stringsAsFactors = FALSE
)

removal_schedules <- list(
  April = month_numbers["Apr"],
  October = month_numbers["Oct"],
  April_October = c(month_numbers["Apr"], month_numbers["Oct"]),
  April_August_November = c(
    month_numbers["Apr"],
    month_numbers["Aug"],
    month_numbers["Nov"]
  )
)

removal_schedule_order <- names(removal_schedules)

paths <- list(
  factors_xlsx = file.path("Inputs", "NH3_Correction_Factors.xlsx"),
  climate_xlsx = file.path("Inputs", "Climate_zone_temperatures.xlsx"),
  output_csv = file.path("Outputs", "Proposed_Tier2_Storage_QC_Monthly.csv"),
  figure_dir = file.path("Outputs", "Storage_QC_Figures")
)

dir.create("Outputs", showWarnings = FALSE, recursive = TRUE)
dir.create(paths$figure_dir, showWarnings = FALSE, recursive = TRUE)

total_months <- simulation_years * length(month_labels)
```

## Read Inputs

```{r}
#| label: read-inputs
#| message: false
#| warning: false

factors <- as.data.frame(
  readxl::read_excel(paths$factors_xlsx, sheet = "Storage"),
  stringsAsFactors = FALSE,
  check.names = FALSE
)

names(factors) <- trimws(names(factors))

text_cols <- c("Animal_Type", "Storage", "Manure_type", "Note")
for (col in intersect(text_cols, names(factors))) {
  factors[[col]] <- trimws(as.character(factors[[col]]))
}
factors$Note[is.na(factors$Note) | factors$Note == "NA"] <- ""

numeric_cols <- c("EFs_TAN", "CT_wet", "CT_dry")
for (col in numeric_cols) {
  factors[[col]] <- suppressWarnings(as.numeric(factors[[col]]))
}

qc_factors <- inner_join(
  qc_combinations,
  factors,
  by = c("Animal_Type", "Storage", "Manure_type")
)

if (nrow(qc_factors) != nrow(qc_combinations)) {
  stop("Could not match all requested QC storage combinations.")
}

climate_raw <- readxl::read_excel(
  paths$climate_xlsx,
  sheet = "Sheet2",
  col_names = FALSE,
  .name_repair = "minimal"
)

month_median_cols <- 2 + (seq_along(month_labels) - 1) * 2

climate_medians <- as.data.frame(
  climate_raw[-c(1, 2), c(1, month_median_cols)],
  stringsAsFactors = FALSE,
  check.names = FALSE
)

names(climate_medians) <- c("Climate_zone", month_labels)
climate_medians$Climate_zone <- trimws(as.character(climate_medians$Climate_zone))

for (month in month_labels) {
  climate_medians[[month]] <- suppressWarnings(as.numeric(climate_medians[[month]]))
}

qc_temperature <- climate_medians %>%
  filter(Climate_zone == qc_climate_zone)

if (nrow(qc_temperature) != 1) {
  stop("Could not find climate zone: ", qc_climate_zone)
}

qc_temperature
qc_factors
```

## Calculation Functions

```{r}
#| label: calculation-functions
#| message: false
#| warning: false

cap_ef <- function(ef_value) {
  pmin(ef_value, ef_max)
}

ct_basis_for_climate <- function(climate_zone) {
  if (grepl("Dry$", climate_zone)) {
    return("dry")
  }

  "wet"
}

calculate_monthly_ef <- function(ef_ref, ct, temperatures_c) {
  if (is.na(ct)) {
    return(cap_ef(rep(ef_ref, length(temperatures_c))))
  }

  cap_ef(ef_ref * ct^(temperatures_c - reference_temperature_c))
}

simulate_storage_monthly <- function(
  factor_row,
  removal_schedule_name,
  removal_months,
  temperatures_c
) {
  ct_basis <- ct_basis_for_climate(qc_climate_zone)
  ct_value <- if (identical(ct_basis, "dry")) factor_row$CT_dry else factor_row$CT_wet

  monthly_ef <- calculate_monthly_ef(
    ef_ref = factor_row$EFs_TAN,
    ct = ct_value,
    temperatures_c = temperatures_c
  )

  stored_tan <- 0
  cumulative_nh3 <- 0
  rows <- vector("list", total_months)

  for (period_index in seq_len(total_months)) {
    month_index <- ((period_index - 1) %% length(month_labels)) + 1
    year_index <- ((period_index - 1) %/% length(month_labels)) + 1

    tan_start <- stored_tan
    tan_removed <- if (month_index %in% removal_months) {
      tan_start * removal_efficiency
    } else {
      0
    }

    tan_after_removal <- tan_start - tan_removed
    available_tan <- tan_after_removal + monthly_tan_kg
    ef_this_month <- monthly_ef[month_index]
    nh3_loss <- min(available_tan, available_tan * ef_this_month)
    monthly_ef_tan_stored <- ifelse(
      available_tan > 0,
      nh3_loss / available_tan,
      NA_real_
    )
    stored_tan <- available_tan - nh3_loss
    cumulative_nh3 <- cumulative_nh3 + nh3_loss

    rows[[period_index]] <- data.frame(
      Animal_Type = factor_row$Animal_Type,
      Storage = factor_row$Storage,
      Manure_type = factor_row$Manure_type,
      Combination = paste(factor_row$Animal_Type, factor_row$Storage, factor_row$Manure_type, sep = " | "),
      Climate_zone = qc_climate_zone,
      Removal_Schedule = removal_schedule_name,
      Removal_Months = paste(month_labels[removal_months], collapse = ";"),
      Year = year_index,
      Month = month_labels[month_index],
      Month_Index = month_index,
      Period = period_index,
      Temperature_C = temperatures_c[month_index],
      CT_Basis = ifelse(identical(ct_basis, "dry"), "Dry", "Wet"),
      CT = ct_value,
      EFs_TAN = factor_row$EFs_TAN,
      Corrected_EF_TAN = ef_this_month,
      TAN_Start_Stored_kg = tan_start,
      TAN_Removed_kg = tan_removed,
      TAN_After_Removal_kg = tan_after_removal,
      TAN_Input_kg = monthly_tan_kg,
      TAN_Available_kg = available_tan,
      NH3_N_kg = nh3_loss,
      Monthly_EF_TAN_Stored = monthly_ef_tan_stored,
      Cumulative_NH3_N_kg = cumulative_nh3,
      TAN_End_Stored_kg = stored_tan,
      stringsAsFactors = FALSE
    )
  }

  do.call(rbind, rows)
}
```

## Monthly QC Data

```{r}
#| label: monthly-qc-data
#| message: false
#| warning: false

temperatures_c <- as.numeric(qc_temperature[month_labels])

qc_rows <- list()
qc_index <- 1

for (row_index in seq_len(nrow(qc_factors))) {
  factor_row <- qc_factors[row_index, ]

  for (schedule_name in names(removal_schedules)) {
    qc_rows[[qc_index]] <- simulate_storage_monthly(
      factor_row = factor_row,
      removal_schedule_name = schedule_name,
      removal_months = removal_schedules[[schedule_name]],
      temperatures_c = temperatures_c
    )
    qc_index <- qc_index + 1
  }
}

qc_monthly <- do.call(rbind, qc_rows) %>%
  mutate(
    Removal_Schedule = factor(Removal_Schedule, levels = removal_schedule_order),
    Period_Label = paste0("Y", Year, "-", Month)
  )

write.csv(
  qc_monthly,
  paths$output_csv,
  row.names = FALSE,
  quote = TRUE,
  na = ""
)

dairy_tank_slurry_monthly <- qc_monthly %>%
  filter(
    Animal_Type == "Dairy Cattle",
    Storage == "Tank",
    Manure_type == "Slurry"
  )

qc_monthly %>%
  select(
    Combination,
    Removal_Schedule,
    Year,
    Month,
    Temperature_C,
    Corrected_EF_TAN,
    TAN_Start_Stored_kg,
    TAN_Removed_kg,
    TAN_Input_kg,
    TAN_Available_kg,
    NH3_N_kg,
    Monthly_EF_TAN_Stored,
    Cumulative_NH3_N_kg,
    TAN_End_Stored_kg
  ) %>%
  mutate(
    across(
      c(
        Temperature_C,
        Corrected_EF_TAN,
        TAN_Start_Stored_kg,
        TAN_Removed_kg,
        TAN_Available_kg,
        NH3_N_kg,
        Monthly_EF_TAN_Stored,
        Cumulative_NH3_N_kg,
        TAN_End_Stored_kg
      ),
      ~ round(.x, 3)
    )
  ) %>%
  knitr::kable()
```

## Monthly TAN In Storage

```{r}
#| label: fig-monthly-tan-storage
#| fig-width: 11
#| fig-height: 6.5

fig_tan_storage <- ggplot(
  qc_monthly,
  aes(x = Period, y = TAN_End_Stored_kg, color = Removal_Schedule)
) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 1.4) +
  facet_wrap(~ Combination, ncol = 1, scales = "free_y") +
  scale_x_continuous(
    breaks = seq(1, total_months, by = 3),
    labels = qc_monthly$Period_Label[seq(1, total_months, by = 3)]
  ) +
  labs(
    title = "Monthly TAN remaining in storage",
    subtitle = paste(qc_climate_zone, "| removal efficiency =", scales::percent(removal_efficiency)),
    x = "Simulation month",
    y = "End-of-month stored TAN, kg",
    color = "Removal schedule"
  ) +
  theme_minimal(base_size = 10) +
  theme(
    panel.grid.minor = element_blank(),
    plot.title = element_text(face = "bold"),
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "bottom"
  )

fig_tan_storage

ggsave(
  file.path(paths$figure_dir, "storage_qc_monthly_tan_in_storage.png"),
  fig_tan_storage,
  width = 11,
  height = 6.5,
  dpi = 300,
  bg = "white"
)
```

## Cumulative NH3-N Over Three Years

```{r}
#| label: fig-cumulative-nh3
#| fig-width: 11
#| fig-height: 6.5

fig_cumulative_nh3 <- ggplot(
  qc_monthly,
  aes(x = Period, y = Cumulative_NH3_N_kg, color = Removal_Schedule)
) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 1.4) +
  facet_wrap(~ Combination, ncol = 1, scales = "free_y") +
  scale_x_continuous(
    breaks = seq(1, total_months, by = 3),
    labels = qc_monthly$Period_Label[seq(1, total_months, by = 3)]
  ) +
  labs(
    title = "Cumulative NH3-N volatilization over three years",
    subtitle = paste(qc_climate_zone, "| monthly TAN input =", monthly_tan_kg, "kg"),
    x = "Simulation month",
    y = "Cumulative NH3-N, kg",
    color = "Removal schedule"
  ) +
  theme_minimal(base_size = 10) +
  theme(
    panel.grid.minor = element_blank(),
    plot.title = element_text(face = "bold"),
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "bottom"
  )

fig_cumulative_nh3

ggsave(
  file.path(paths$figure_dir, "storage_qc_cumulative_nh3.png"),
  fig_cumulative_nh3,
  width = 11,
  height = 6.5,
  dpi = 300,
  bg = "white"
)
```

## Year 3 Summary

```{r}
#| label: year-3-summary

year3_summary <- qc_monthly %>%
  filter(Year == simulation_years) %>%
  group_by(Combination, Removal_Schedule, Climate_zone) %>%
  summarize(
    Year3_TAN_Input_kg = sum(TAN_Input_kg),
    Year3_TAN_Available_kg = sum(TAN_Available_kg),
    Year3_NH3_N_kg = sum(NH3_N_kg),
    Year3_EF_TAN = weighted.mean(
      Monthly_EF_TAN_Stored,
      TAN_Available_kg,
      na.rm = TRUE
    ),
    Year3_Unweighted_Monthly_EF_TAN = mean(
      Monthly_EF_TAN_Stored,
      na.rm = TRUE
    ),
    Year3_Mass_Balance_EF_TAN = Year3_NH3_N_kg / Year3_TAN_Input_kg,
    Final_Stored_TAN_kg = dplyr::last(TAN_End_Stored_kg),
    .groups = "drop"
  ) %>%
  mutate(
    Year3_NH3_N_kg = round(Year3_NH3_N_kg, 3),
    Year3_EF_TAN = round(Year3_EF_TAN, 3),
    Year3_Unweighted_Monthly_EF_TAN = round(
      Year3_Unweighted_Monthly_EF_TAN,
      3
    ),
    Year3_Mass_Balance_EF_TAN = round(Year3_Mass_Balance_EF_TAN, 3),
    Final_Stored_TAN_kg = round(Final_Stored_TAN_kg, 3)
  )

knitr::kable(year3_summary)
```

## Output Files

```{r}
#| label: output-files

output_files <- data.frame(
  File = c(
    paths$output_csv,
    file.path(paths$figure_dir, "storage_qc_monthly_tan_in_storage.png"),
    file.path(paths$figure_dir, "storage_qc_cumulative_nh3.png")
  ),
  stringsAsFactors = FALSE
)

knitr::kable(output_files)
```
