---
title: "Proposed Tier 2 Housing NH3 Emission Factors"
format: html
---

## Overview

This file calculates proposed Tier 2 housing ammonia emission factors by
animal type, housing type, manure type, and broad climate zone.

The deterministic output uses annual temperature and the updated housing
correction-factor table:

```text
EF = EFs_TotalN * CT_15^(T_annual - 15)
```

`EFs_TotalN` is treated as the reference EF at 15 degrees C for this proposal,
even though the source table describes it as a global average.

## Preparation

```{r}
#| label: preparation
#| message: false
#| warning: false

library(readxl)

# Month order used in the climate workbook.
month_labels <- c(
  "Jan", "Feb", "Mar", "Apr", "May", "Jun",
  "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"
)

# Upper cap for emission factors.
# Revise this value if a different maximum EF is needed.
ef_max <- 1

# Monte Carlo settings.
set.seed(2026)
n_iter <- 5000

# Project file paths.
paths <- list(
  factors_xlsx = file.path("Inputs", "NH3_Correction_Factors.xlsx"),
  climate_xlsx = file.path("Inputs", "Climate_zone_temperatures.xlsx"),
  proposed_csv = file.path("Outputs", "Proposed_Tier2_Housing.csv"),
  mc_summary_csv = file.path("Outputs", "Proposed_Tier2_Housing_MC_Summary.csv")
)

# The final proposal uses four broad climate zones.
broad_climate_map <- list(
  Tropical = c("Tropical Wet", "Tropical Moist", "Tropical Dry"),
  Warm_Temperate = c("Warm Temperate Moist", "Warm Temperate Dry"),
  Cool_Temperate = c("Cool Temperate Moist", "Cool Temperate Dry"),
  Boreal = c("Boreal Moist", "Boreal Dry")
)

id_cols <- c("Animal_Type", "Housing", "Manure_type")
climate_cols <- names(broad_climate_map)
```

## Input Data

```{r}
#| label: input-data
#| message: false
#| warning: false

# Read the updated housing EF and CT table.
factors <- as.data.frame(
  readxl::read_excel(paths$factors_xlsx, sheet = "Housing"),
  stringsAsFactors = FALSE,
  check.names = FALSE
)

names(factors) <- trimws(names(factors))

# Keep the text columns stable for matching and output.
text_cols <- c(id_cols, "Note")
for (col in intersect(text_cols, names(factors))) {
  factors[[col]] <- trimws(as.character(factors[[col]]))
}
factors$Note[is.na(factors$Note) | factors$Note == "NA"] <- ""

# Convert numeric columns from Excel to numeric R vectors.
numeric_cols <- c(
  "EFs_TotalN", "EFs_Lower", "EFs_Upper",
  "EF_Ref_15", "CT_15", "CT_Lower", "CT_Upper"
)
for (col in numeric_cols) {
  factors[[col]] <- suppressWarnings(as.numeric(factors[[col]]))
}

# Recreate the output skeleton directly from the updated input sheet.
proposed_result <- factors[id_cols]
```

## Climate Temperatures

```{r}
#| label: climate-temperatures
#| message: false
#| warning: false

# Sheet2 has two header rows: month names in row 1 and Median/Std in row 2.
climate_raw <- readxl::read_excel(
  paths$climate_xlsx,
  sheet = "Sheet2",
  col_names = FALSE,
  .name_repair = "minimal"
)

month_median_cols <- 2 + (seq_along(month_labels) - 1) * 2
month_std_cols <- 3 + (seq_along(month_labels) - 1) * 2

climate_medians <- as.data.frame(
  climate_raw[-c(1, 2), c(1, month_median_cols)],
  stringsAsFactors = FALSE,
  check.names = FALSE
)
climate_stds <- as.data.frame(
  climate_raw[-c(1, 2), c(1, month_std_cols)],
  stringsAsFactors = FALSE,
  check.names = FALSE
)

names(climate_medians) <- c("Climate_zone", month_labels)
names(climate_stds) <- c("Climate_zone", month_labels)

clean_climate_table <- function(climate_table) {
  climate_table$Climate_zone <- trimws(as.character(climate_table$Climate_zone))
  climate_table <- climate_table[
    !is.na(climate_table$Climate_zone) &
      nzchar(climate_table$Climate_zone) &
      climate_table$Climate_zone != "0",
    ,
    drop = FALSE
  ]

  for (month in month_labels) {
    climate_table[[month]] <- suppressWarnings(as.numeric(climate_table[[month]]))
  }

  climate_table
}

climate_medians <- clean_climate_table(climate_medians)
climate_stds <- clean_climate_table(climate_stds)

get_climate_matrix <- function(climate_table, component_zones) {
  climate_rows <- climate_table[climate_table$Climate_zone %in% component_zones, ]

  if (nrow(climate_rows) != length(component_zones)) {
    stop(
      "Expected climate rows for: ",
      paste(component_zones, collapse = ", ")
    )
  }

  as.matrix(climate_rows[month_labels])
}

# Annual temperatures are calculated after aggregating component climate zones
# into the four broad zones.
broad_climate_summary <- data.frame(
  Climate_zone = climate_cols,
  Annual_Temperature_C = NA_real_,
  Annual_Temperature_SD_C = NA_real_,
  stringsAsFactors = FALSE
)

for (climate_col in climate_cols) {
  component_zones <- broad_climate_map[[climate_col]]
  median_matrix <- get_climate_matrix(climate_medians, component_zones)
  std_matrix <- get_climate_matrix(climate_stds, component_zones)

  monthly_broad_medians <- colMeans(median_matrix, na.rm = TRUE)

  broad_climate_summary[
    broad_climate_summary$Climate_zone == climate_col,
    "Annual_Temperature_C"
  ] <- mean(monthly_broad_medians, na.rm = TRUE)

  # The Monte Carlo uses one annual temperature draw. Its SD is the simple
  # average of all component-zone monthly Std values for the broad climate zone.
  broad_climate_summary[
    broad_climate_summary$Climate_zone == climate_col,
    "Annual_Temperature_SD_C"
  ] <- mean(as.numeric(std_matrix), na.rm = TRUE)
}

broad_climate_summary
```

## Calculation Functions

```{r}
#| label: calculation-functions
#| message: false
#| warning: false

cap_ef <- function(ef_value) {
  pmin(ef_value, ef_max)
}

# Notes such as "Only applicable to T >15" define the minimum annual
# temperature required before CT is applied.
threshold_from_note <- function(note) {
  note <- as.character(note)
  match_text <- regmatches(
    note,
    regexpr(">[[:space:]]*[-+]?[0-9]*\\.?[0-9]+", note)
  )

  if (length(match_text) == 0 || is.na(match_text) || match_text == "") {
    return(NA_real_)
  }

  as.numeric(gsub(">|[[:space:]]", "", match_text))
}

ct_is_applicable <- function(ct, annual_temperature_c, note) {
  if (is.na(ct)) {
    return(FALSE)
  }

  threshold_c <- threshold_from_note(note)
  if (is.na(threshold_c)) {
    return(TRUE)
  }

  annual_temperature_c > threshold_c
}

calculate_annual_ef <- function(ef_ref, ct, annual_temperature_c, note) {
  if (ct_is_applicable(ct, annual_temperature_c, note)) {
    corrected_ef <- ef_ref * ct^(annual_temperature_c - 15)
  } else {
    corrected_ef <- ef_ref
  }

  cap_ef(corrected_ef)
}
```

## Annual EF Calculation

```{r}
#| label: annual-ef-calculation
#| message: false
#| warning: false

deterministic_detail_rows <- list()
deterministic_detail_index <- 1

for (row_index in seq_len(nrow(factors))) {
  for (climate_col in climate_cols) {
    annual_temp <- broad_climate_summary[
      broad_climate_summary$Climate_zone == climate_col,
      "Annual_Temperature_C"
    ]

    ef_value <- calculate_annual_ef(
      ef_ref = factors$EFs_TotalN[row_index],
      ct = factors$CT_15[row_index],
      annual_temperature_c = annual_temp,
      note = factors$Note[row_index]
    )

    proposed_result[row_index, climate_col] <- round(ef_value, 3)

    deterministic_detail_rows[[deterministic_detail_index]] <- data.frame(
      Animal_Type = factors$Animal_Type[row_index],
      Housing = factors$Housing[row_index],
      Manure_type = factors$Manure_type[row_index],
      Climate_zone = climate_col,
      Annual_Temperature_C = round(annual_temp, 3),
      EFref_TotalN = factors$EFs_TotalN[row_index],
      CT_15 = factors$CT_15[row_index],
      CT_Threshold_C = threshold_from_note(factors$Note[row_index]),
      CT_Applied = ct_is_applicable(
        factors$CT_15[row_index],
        annual_temp,
        factors$Note[row_index]
      ),
      EF_TotalN = round(ef_value, 3),
      stringsAsFactors = FALSE
    )

    deterministic_detail_index <- deterministic_detail_index + 1
  }
}

deterministic_detail <- do.call(rbind, deterministic_detail_rows)
```

## Monte Carlo Preparation

```{r}
#| label: monte-carlo-preparation
#| message: false
#| warning: false

# Bounded lognormal sampling uses the reported lower and upper 95% CI limits
# as bounds. Values outside the bounds are resampled.
sample_bounded_lognormal <- function(n, center, lower, upper) {
  if (is.na(center) || is.na(lower) || is.na(upper) ||
      center <= 0 || lower <= 0 || upper <= 0 || lower >= upper) {
    return(rep(center, n))
  }

  sdlog <- (log(upper) - log(lower)) / (2 * 1.96)
  samples <- rep(NA_real_, n)
  remaining <- seq_len(n)
  attempts <- 0

  while (length(remaining) > 0 && attempts < 100) {
    draws <- rlnorm(length(remaining), meanlog = log(center), sdlog = sdlog)
    keep <- draws >= lower & draws <= upper

    if (any(keep)) {
      samples[remaining[keep]] <- draws[keep]
    }

    remaining <- remaining[!keep]
    attempts <- attempts + 1
  }

  # If rejection sampling leaves a few empty values, use the point estimate.
  samples[is.na(samples)] <- center
  samples
}

temperature_samples <- list()
for (climate_col in climate_cols) {
  annual_temp <- broad_climate_summary[
    broad_climate_summary$Climate_zone == climate_col,
    "Annual_Temperature_C"
  ]
  annual_sd <- broad_climate_summary[
    broad_climate_summary$Climate_zone == climate_col,
    "Annual_Temperature_SD_C"
  ]

  temperature_samples[[climate_col]] <- rnorm(
    n_iter,
    mean = annual_temp,
    sd = annual_sd
  )
}
```

## Monte Carlo Simulation

```{r}
#| label: monte-carlo-simulation
#| message: false
#| warning: false

mc_summary_rows <- list()
mc_summary_index <- 1

for (row_index in seq_len(nrow(factors))) {
  ef_samples <- sample_bounded_lognormal(
    n = n_iter,
    center = factors$EFs_TotalN[row_index],
    lower = factors$EFs_Lower[row_index],
    upper = factors$EFs_Upper[row_index]
  )

  ct_samples <- sample_bounded_lognormal(
    n = n_iter,
    center = factors$CT_15[row_index],
    lower = factors$CT_Lower[row_index],
    upper = factors$CT_Upper[row_index]
  )

  threshold_c <- threshold_from_note(factors$Note[row_index])

  for (climate_col in climate_cols) {
    temp_samples <- temperature_samples[[climate_col]]

    if (is.na(threshold_c)) {
      ct_applied <- !is.na(ct_samples)
    } else {
      ct_applied <- !is.na(ct_samples) & temp_samples > threshold_c
    }

    ef_sim <- ifelse(
      ct_applied,
      ef_samples * ct_samples^(temp_samples - 15),
      ef_samples
    )
    ef_sim <- cap_ef(ef_sim)

    deterministic_ef <- proposed_result[row_index, climate_col]

    mc_summary_rows[[mc_summary_index]] <- data.frame(
      Animal_Type = factors$Animal_Type[row_index],
      Housing = factors$Housing[row_index],
      Manure_type = factors$Manure_type[row_index],
      Climate_zone = climate_col,
      Annual_Temperature_C = round(
        broad_climate_summary[
          broad_climate_summary$Climate_zone == climate_col,
          "Annual_Temperature_C"
        ],
        3
      ),
      Annual_Temperature_SD_C = round(
        broad_climate_summary[
          broad_climate_summary$Climate_zone == climate_col,
          "Annual_Temperature_SD_C"
        ],
        3
      ),
      EF_Deterministic = round(deterministic_ef, 3),
      EF_MC_Mean = round(mean(ef_sim, na.rm = TRUE), 3),
      EF_MC_Lower95 = round(as.numeric(quantile(ef_sim, 0.025, na.rm = TRUE)), 3),
      EF_MC_Upper95 = round(as.numeric(quantile(ef_sim, 0.975, na.rm = TRUE)), 3),
      CT_Threshold_C = threshold_c,
      stringsAsFactors = FALSE
    )

    mc_summary_index <- mc_summary_index + 1
  }
}

mc_summary <- do.call(rbind, mc_summary_rows)
```

## Monte Carlo Summary

```{r}
#| label: monte-carlo-summary
#| message: false
#| warning: false

head(mc_summary)
```

## Validation

```{r}
#| label: validation
#| message: false
#| warning: false

expected_output_cols <- c(id_cols, climate_cols)

if (!identical(names(proposed_result), expected_output_cols)) {
  stop("Proposed output columns changed unexpectedly.")
}

if (nrow(proposed_result) != nrow(factors)) {
  stop("Proposed output row count does not match the Housing sheet.")
}

if (nrow(mc_summary) != nrow(factors) * length(climate_cols)) {
  stop("Monte Carlo summary row count changed unexpectedly.")
}

if (any(is.na(proposed_result[climate_cols]))) {
  stop("Proposed climate-zone EF columns should not contain missing values.")
}

if (max(as.matrix(proposed_result[climate_cols]), na.rm = TRUE) > ef_max) {
  stop("Proposed output exceeds ef_max.")
}

mc_ef_cols <- c("EF_Deterministic", "EF_MC_Mean", "EF_MC_Lower95", "EF_MC_Upper95")
if (max(as.matrix(mc_summary[mc_ef_cols]), na.rm = TRUE) > ef_max) {
  stop("Monte Carlo summary exceeds ef_max.")
}
```

## Write Outputs

```{r}
#| label: write-outputs
#| message: false
#| warning: false

write.csv(
  proposed_result,
  paths$proposed_csv,
  row.names = FALSE,
  quote = TRUE,
  na = ""
)

write.csv(
  mc_summary,
  paths$mc_summary_csv,
  row.names = FALSE,
  quote = TRUE,
  na = ""
)
```

## Summary

```{r}
#| label: summary
#| message: false
#| warning: false

output_summary <- data.frame(
  output_file = c(paths$proposed_csv, paths$mc_summary_csv),
  rows = c(nrow(proposed_result), nrow(mc_summary)),
  columns = c(ncol(proposed_result), ncol(mc_summary)),
  stringsAsFactors = FALSE
)

output_summary
```

## Method Notes

The final proposal uses annual temperature only. Monthly temperature medians
are used only to construct annual broad-climate temperatures from the climate
workbook.

For the Monte Carlo simulation, monthly temperatures are not sampled
independently. Instead, one annual temperature is sampled for each broad
climate zone using the annual mean temperature and an annual-level standard
deviation derived from the workbook's monthly `Std` values. This avoids
unrealistic simulated years where adjacent months move in unrelated directions.

The correction uses `EFs_TotalN` as the 15 degrees C reference EF. This follows
the requested output method, but it is an assumption because `EFs_TotalN` is a
global average and the source data do not identify the average observation
temperature for that EF. The input column `EF_Ref_15` is retained in the input
table but is not used in the main proposed output.
